Project eda

### 2024-03-19 R/eda-code/eda.qmd
eda map

---
title: "An example exploratory analysis script"
author: "Andreas Handel"
date: "10/28/2021"
output: html_document
---

@@

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

# Data exploration through tables

Showing a bit of code to produce and save a summary table.

```{r}
summary_df = skimr::skim(mydata)
print(summary_df)
# save to file
summarytable_file = here("results", "tables", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```

linear model applied

```{r}
# Fit the ANOVA model
model <- aov(Selective ~ Seed + Strain + Treatment + Day + Rep, data = mydata)

# Summarize the ANOVA results
linear_model <- summary(model)

# Save this table

anovatable_file = here("results", "tables", "anovatable.rds")
saveRDS(linear_model, file = anovatable_file)
```

>>>>

---
title: "Project eda"
author: "Xueyan Hu"
date: "2023-02-22"
output: html_document
---

@@

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata_cancer.rds")
#load data
mydata <- readRDS(data_location)
```

# bar chart for crude prevalence according to states

```{r}
# Group by state and calculate the total population and cancer cases
state_total <- mydata %>%
  group_by(StateDesc) %>%
  summarize(TotalPopulation = sum(PopulationCount),
            TotalCancerCases = sum(Cancer_Case))

# Calculate the prevalence for each state
state_total <- state_total %>%
  mutate(TotalPrevalence = TotalCancerCases / TotalPopulation)

# View the resulting dataframe
print(state_total)

# Load usmap package and get the state data
library(usmap)

usmap_state_data <- us_map(regions = "states")

# There are a few state names that are cut off in state_total, but the column
# is already alphabetical so we can replace it.
# To use plot_usmap() we have to have a column named `state` or `fips`.
state_total$state <- usmap_state_data$full
print(state_total)

# Now we can try to make a usmap plot
plot_usmap("states", data = state_total, values = "TotalPrevalence") +
  scale_fill_viridis_c()
```

# List 10 cities that has the highest cancer prevalence

```{r}
top_10_highest <- mydata %>%
  top_n(10, Data_Value) %>%
  select(Year, StateAbbr, CityName, Data_Value)

print(top_10_highest)

# create a table
table <- top_10_highest %>%
  gt() %>%
  tab_spanner(
    label = md('**Year**'),
    columns = c('Year')
  ) %>%
  tab_spanner(
    label = md('**State**'),
    columns = c('StateAbbr')
  ) %>%
  tab_spanner(
    label =  md('**Cancer Prevalence**'),
    columns = c('Data_Value')
  ) %>%
  tab_header(
    title = '2017 top 10 cities which have highest cancer rates'
  ) %>%
  opt_stylize(style = 6, color = 'gray')
```

We are saving the results to the `results` folder. Depending on how many tables/figures you have, it might make sense to have separate folders for each. And/or you could have separate folders for exploratory tables/figures and for final tables/figures. Just choose a setup that makes sense for your project and works for you, and provide enough documentation that someone can understand what you are doing.

# fit a simple linear model

Before we started analysis, we had an assumption

```{r}
# Fit linear regression model
lm1 <- lm(Data_Value ~ PopulationCount, data = mydata)

# Print summary of the model
summary(lm1)
```
So the hypothesis is not supported by the linear model fitting result. There is no obvious evidence that the cancer prevalence in larger cities (higher population)
